---
title: "How to add your dataset?"
output: md_document
date: "2022-09-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Creation of a Dataset folder

Each dataset is put into a folder called "Names_Year" for example "Serbin_et_al_2019". We recommand to put into the folder the article associated with the dataset or the protocol of measurement, which should include the description of the gas exchange, spectrometers, and of the species. 

## Description of the dataset

In each dataset folder a csv file called Description.csv has to be included. An example is given in the Folder 0_Template. The latitude and longitude coordinates will be used to position the dataset on a world map.

```{r}
Description=read.csv(file='Description.csv')
knitr::kable(Description, "html")
```

## Adding the gas exchange A-Ci data to the dataset folder

The idea of this project is to include the raw A-Ci data so we can fit the curves and estimate Vcmax using the same method for all datasets.

We have requirements for the final A-Ci data to be used by the fitting procedure (see table below with the needed column names). However, we don't have hard requirements in the way to obtain this final data. 

The A-Ci data should be cleaned from spurious measurements and points that would impact Vcmax or Jmax estimation should not be included. Note that the SampleID will be used by the fitting procedure to identify individual A-Ci curves. If you made several curves for the same leaf we recommend to only keep the best one. 



```{r}
Description=read.csv(file='Aci_data.csv')
knitr::kable(Description, "html")
```
